Quantizing the Bosonic String on a Loop Quantum Gravity Background

### 2024-06-27 lqg-strings.tex
Completed holography section

First let us recall the form of the Polyakov action for the bosonic string \eqref{eqn:polyakov-action}:
\begin{equation*}
	S_{P} = -\frac{T}{2} \int \diff \tau \wedge \diff \sigma \sqrt{-g} g^{ab} \partial_a X^\mu \partial_b X^\nu \eta_{\mu\nu}.
\end{equation*}
If one was to ask a beginning graduate student as to how one can incorporate a gauge field in the above expression, the natural answer would be that we can promote the $X^\mu$ to vectors which live in the representation of some gauge group $\mc G$ and promote the ordinary derivative to a gauge covariant derivative \wrt a connection valued in $\mc G$. The resulting action would now have the form:
\begin{equation}\label{eqn:polyakov-gauged}
	S_{P} = -\frac{T}{2} \int \diff \tau \wedge \diff \sigma \sqrt{-g} g^{ab} \mc{D}_a X^\mu \mc{D}_b X^\nu \eta_{\mu\nu},
\end{equation}
where the covariant derivative $\mc{D}_a$ is of the form:
\begin{equation}\label{eqn:covariant-deriv-v1}
	\mc{D}_a X^\mu(\tau , \sigma) = \partial_a X^\mu + k \mc{A}_{a}^\mu{}_\nu X^\nu,
\end{equation}
where $\mc{A}_{a}^\mu{}_\nu$ is the gauge field and $k$ is the gauge coupling. A more experienced researcher would object saying that we know that the $X^\mu$ are spacetime co-ordinates and not vectors transforming under the gauge group $\mc G$. But we will see that this is not a problem. In fact, in order to be able to couple the string to an \emph{arbitrarily} curved geometry we will find that we \emph{require} the $X^\mu$ to transform as vectors under the action of some \emph{local} gauge field.

@@

Strictly speaking the above expression corresponds to an $\mf{iso}(3,1)$ Lie algebra generator only when $\ep = 0$. $l$ is a length scale which must be introduced so that the components $(p^0, p^i)$ can be viewed as components of a ``translation vector'' living in the tangent space of the $3+1$ dimensional geometry $\mb{\Gamma}$. Similarly the boosts and rotations generators are for objects living in $\mb{\Gamma}$.

Let us write down the gauged Polyakov action for the Poincaré lie algebra valued connection $\mc{A}_a$ whose components are given in \eqref{eqn:poincare-lie-algeba}:
\begin{equation}\label{eqn:poincare-polyakov}
	S'_{P} = -\frac{T}{2} \int \diff^2 x \sqrt{-g} g^{ab} \mc{D}_a X^A \mc{D}_b X^B \eta_{AB},
\end{equation}
where now the $X^A = \{X^\mu, X^r\}$ (as mentioned previously in \eqref{eqn:enlarged-Xs}) and $\eta_{AB}$ is the flat metric on the $d+2$ enlarged bulk manifold $\mb{M}$ and the covariant derivative is of the form defined in \eqref{eqn:covariant-deriv-v1}, but only now acting on $X^A$ instead of on $X^\mu$:
\begin{equation}\label{eqn:covariant-deriv-v2}
	\mc{D}_a X^A(\tau , \sigma) = \partial_a X^A + k \mc{A}_{a}^A{}_B X^B.
\end{equation}
Now, we will expand out the action \eqref{eqn:poincare-polyakov} and express it in terms of the $X^\mu$ and $X^r$ separately. It is clear that the resulting expression will be sum of three terms:
\begin{align}
	S'_{P} & = -\frac{T}{2} \int \diff^2 x \sqrt{-g} g^{ab} \left( \mc{D}_a X^\mu \mc{D}_b X^\nu \eta_{\mu\nu} + \right. \nonumber \\
	& \left. \mc{D}_a X^\mu \mc{D}_b X^r \eta_{\mu r} + \mc{D}_a X^r \mc{D}_b X^r \eta_{rr} \right)
\end{align}

\textbf{todo:}

\begin{itemize}
	\item Explain why deSitter, anti-deSitter and flat space are the only consistent solutions for the enlarged bulk geometry.
	\item Recall once again the steps which led us here.
\end{itemize}

>>>>

First let us recall the form of the Polyakov action for the bosonic string \eqref{eqn:polyakov-action}:
\begin{equation*}
	S_{P} = -\frac{T}{2} \int \diff \tau \wedge \diff \sigma \sqrt{-g} g^{ab} \partial_a X^\mu \partial_b X^\nu \eta_{\mu\nu}.
\end{equation*}
If one was to ask a beginning graduate student as to how one can incorporate a gauge field in the above expression, the natural answer would be that we can promote the $X^\mu$ to vectors which live in the representation of some gauge group $\mc G$ and promote the ordinary derivative to a gauge covariant derivative \wrt a connection valued in $\mc G$. The resulting action would now have the form:
\begin{equation}\label{eqn:polyakov-gauged}
	S_{P} = -\frac{T}{2} \int \diff \tau \wedge \diff \sigma \sqrt{-g} g^{ab} D_a X^\mu D_b X^\nu \eta_{\mu\nu},
\end{equation}
where the covariant derivative $D_a$ is of the form:
\begin{equation}\label{eqn:covariant-deriv-v1}
	D_a X^\mu(\tau , \sigma) = \partial_a X^\mu + k \mc{A}_{a}^\mu{}_\nu X^\nu,
\end{equation}
where $\mc{A}_{a}^\mu{}_\nu$ is the gauge field and $k$ is the gauge coupling. A more experienced researcher would object saying that we know that the $X^\mu$ are spacetime co-ordinates and not vectors transforming under the gauge group $\mc G$. But we will see that this is not a problem. In fact, in order to be able to couple the string to an \emph{arbitrarily} curved geometry we will find that we \emph{require} the $X^\mu$ to transform as vectors under the action of some \emph{local} gauge field.

@@

Strictly speaking the above expression corresponds to an $\mf{iso}(3,1)$ Lie algebra generator only when $\ep = 0$. The case when $\ep = \pm 1$ will be motivated and discussed later. $l$ is a length scale which must be introduced so that the components $(p^0, p^i)$ can be viewed as components of a ``translation vector'' living in the tangent space of the $3+1$ dimensional geometry $\mb{\Gamma}$. Similarly the boosts and rotations generators are for objects living in $\mb{\Gamma}$.

Let us write down the gauged Polyakov action for the Poincaré lie algebra valued connection $\mc{A}_a$ whose components are given in \eqref{eqn:poincare-lie-algeba}:
\begin{equation}\label{eqn:poincare-polyakov}
	S'_{P} = -\frac{T}{2} \int \diff^2 x \sqrt{-g} g^{ab} \mc{D}_a X^A \mc{D}_b X^B \eta_{AB},
\end{equation}
where now the $X^A = \{X^\mu, X^r\}$ (as mentioned previously in \eqref{eqn:enlarged-Xs}) and $\eta_{AB}$ is the flat metric on the $d+2$ enlarged bulk manifold $\mb{M}$ and the covariant derivative is of the form defined in \eqref{eqn:covariant-deriv-v1}, but only now acting on $X^A$ instead of on $X^\mu$:
\begin{equation}\label{eqn:covariant-deriv-v2}
	\mc{D}_a X^A(\tau , \sigma) = \partial_a X^A + k \mc{A}_{a}^A{}_B X^B.
\end{equation}
Now, we will expand out the action \eqref{eqn:poincare-polyakov} and express it in terms of the $X^\mu$ and $X^r$ separately. It is clear that the resulting expression will be sum of three terms, with the first term being the (gauged) Polyakov action for the string on $\mb{\Gamma}$, terms describing the coupling of the string on $\mb{\Gamma}$ with fluctuations along the extra dimensions $X^r$ and finally a Polyakov like term for the lone extra dimension. We can write this schematically as:
\begin{equation}\label{eqn:poincare-polyakov-decomp}
	S'_P[X^A] = S_P[X^\mu] + S_{int}[X^\mu, X^r] + S_P[X^r].
\end{equation}
One can now argue along the same lines as the original arguments which led to the discovery of D-branes \cite{Dai1989New-Connections,Polchinski1995Dirichlet-Branes,Horava1989Background} to say that the extra dimension $X^r$ will become dynamical in the full theory even though it began life only as a classical constraint on the string worldsheet's location. This will imply that while most of the time the string will remain confined to $\mb{\Gamma}$, fluctuations along $X^r$ can cause the string to leave the ``brane'' and venture out into the bulk $\mb{M}$. If this is truly the case then the geometry of the enlarged bulk $\mb{M}$ will be essential for a description of the full string dynamics.

This is where the non-zero value of $\ep$ mentioned in \eqref{eqn:poincare-lie-algeba} come into play. It turns out that for $\ep = +1$ ($\ep = -1$) the connection in \eqref{eqn:poincare-lie-algeba} describes a $\mf{so}(4,1)$ ($\mf{so}(3,2)$) connection respectively. These describe the symmetries of (asymptotically) deSitter and anti-deSitter spacetimes respectively. The length scale $l$ determines the value of the effective cosmological constant in the dS (adS) geometries:
\begin{equation}\label{eqn:cosmological-constant}
	\lambda = \frac{3 \ep}{l^2}.
\end{equation}
One might ask what is the motivation for introducing dS (adS) geometries into the discussion. The main motivation of course is the Maldacena or adS/CFT correspondence. The adS spacetime is usually understood to arise as a consequence of non-perturbative effects which nobody really has a proper understanding of as yet. One might hope that for a correspondence which has passed all possible checks uptil now, there should be some way to realise it in the classical theory also.

What our discussion in this section shows, however, is that such a classical description is possible if we switch from the metric based approach to the connection based one for describing the background geometry and its coupling to the string worldsheet.

% \begin{align}
% 	S'_{P} & = -\frac{T}{2} \int \diff^2 x \sqrt{-g} g^{ab} \left( \mc{D}_a X^\mu \mc{D}_b X^\nu \eta_{\mu\nu} + \right. \nonumber \\
% 	& \left. \mc{D}_a X^\mu \mc{D}_b X^r \eta_{\mu r} + \mc{D}_a X^r \mc{D}_b X^r \eta_{rr} \right)
% \end{align}
